---
title: "Amit's Reading List"
subtitle: "A lifetime is not enough to read all the books I want to read!" 
date: last-modified
author: "Amit Arora"
slug: index
theme: cosmo
---

```{r setup}
#| echo: false
#| warning: false
library(gt)
library(glue)
library(dplyr)
library(tidyr)
library(scales)
library(ggtext)
library(ggpmisc)
library(ggthemes)
library(gtExtras)
library(janitor)
library(tidyverse)

books <- read_csv("books.csv") %>%
    janitor::clean_names()
books_read <- books %>%
  filter(reading_status == "Finished")
genres <- unique(books$genre)

total <- nrow(books)
total_adjusted <- total - nrow(books %>% filter(reading_status == "On the bookshelf, but probably not gonna read" | reading_status == "Started but could not finish"))
finished <- nrow(books %>% filter(reading_status == "Finished"))
reading_backlog <- round((total_adjusted - finished)/total_adjusted, 2)

reading_backlog_status <- case_when(
  reading_backlog < 0.3 ~ "a very healthy",
  reading_backlog >= 0.3 & reading_backlog < 0.5 ~ "a reasonably ok",
  TRUE ~ "an unhealthy"
)
# print(genres)
# print(unique(books$reading_status))
```
As a child I liked reading books but then over the years I lost the habit. Then a friend gifted me a book on my birthday and it helped me rediscover reading. There were two other factors that helped me get back to reading:

1. The Covid lockdowns of 2020 provided a little more time at hand. I could give time and energy to things I really wanted to do.

1. One good  habit begets another, so as I started focussing more on physical fitness it had other positive side effects. I thought what else could I accomplish that I never thought possible, getting back to reading was a long standing desire.

To track the effectiveness of my reading habit, I define a metric that I like to call `reading backlog`. The `reading backlog` is defined as follows:
$$
\frac{number\:of\:books\:to\:read}{number\:of\:books\:excluding\:the\:ones\:I\:would\:probably\:never\:read}
$$

The lower the `reading backlog` value the better. As of `r Sys.Date()`, the `reading backlog` stands at _`r reading_backlog_status`_ value of `r reading_backlog`. The number of books on my bookshelf that I still plan to read is `r total_adjusted - finished`.

Here is a graphic showing what genres I like to read. Productiviy and health stand out.

```{r}
#| echo: false
#| warning: false
book_counts_by_genre <- books %>%
  #filter(reading_status == "Finished") %>%
  count(genre, sort=TRUE)

genre_w_most_books1 <- book_counts_by_genre$genre[1]
genre_w_most_books2 <- book_counts_by_genre$genre[2]
genre_w_most_books3 <- book_counts_by_genre$genre[3]

book_counts_by_genre %>%
  ggplot(aes(x=reorder(genre, -n), y=n, col=genre, fill=genre)) + 
  geom_bar(stat="identity") + 
  scale_y_continuous(breaks = pretty_breaks()) + 
  coord_flip() +
  scale_color_tableau(palette = "Tableau 20",) + 
  scale_fill_tableau(palette = "Tableau 20") +
  labs(title="I like to read books across different genres",
       subtitle = glue("<span style='color:#FF00FF'>**\"{genre_w_most_books1}\", \"{genre_w_most_books2}\" and \"{genre_w_most_books3}\"**</span> are the genres I like the most!")) +
  theme_fivethirtyeight() +
  theme(legend.position = "none", plot.subtitle = element_markdown()) +
  theme(axis.title = element_text()) +
  ylab('Number of books') +
  xlab("")
```

```{r}
#| execute: false
#| echo: false
#| warning: false
p <- books %>%
  filter(reading_status == "Finished") %>%
  count(year_read_or_purchased, sort=TRUE) %>%
  ggplot(aes(x=year_read_or_purchased, y=n)) +
  geom_line(size=1) + 
  geom_point(size=2) +
  scale_y_continuous(labels = comma) +
  theme_fivethirtyeight() +
  scale_color_tableau(palette="Tableau 20") +
  scale_fill_tableau(palette="Tableau 20") +
  labs(title="Books read per year") +
  theme(legend.position="bottom") +
  theme(text = element_text(size=15)) +
  theme(axis.title = element_text()) +
  ylab("Number of books read") + xlab("")
```
```{r}
#| echo: false
#| warning: false
books %>%
  replace_na(list(notes = '-')) %>%
  arrange(desc(year_read_or_purchased)) %>%
  gt() %>%
  gt_highlight_rows(rows=reading_status=="Currently Reading", font_weight = "normal") %>%
  tab_style(
    style = list(
      cell_fill(color = "darkolivegreen1")
    ),
    locations = cells_body(
      
      rows = reading_status=="Finished")
  ) %>%
  tab_style(
    style = list(
      cell_fill(color = "lightcyan")
    ),
    locations = cells_body(
      
      rows = reading_status=="Partially Read")
  ) %>%
  cols_label(
    book = "Book",
    author = "Author",
    genre = "Genre",
    year_read_or_purchased = "Year",
    reading_status="Reading Status",
    notes="Notes"
  ) %>%
  tab_header(title="Amit's book list",
             subtitle=glue("{nrow(books)} books, {length(genres)} genres, {max(books$year_read_or_purchased, na.rm=TRUE)-min(books$year_read_or_purchased, na.rm=TRUE)} years")) %>%
  tab_style(
     locations = cells_column_labels(columns = everything()),
     style     = list(
       #Give a thick border below
       cell_borders(sides = "bottom", weight = px(3)),
       #Make text bold
       cell_text(weight = "bold")
     )
   ) %>% 
   #Apply different style to the title
   tab_style(
     locations = cells_title(groups = "title"),
     style     = list(
       cell_text(weight = "bold", size = 24)
     )
   )
```

:::{.callout-note}
This website was created using Quarto and GitHub actions, please see [https://github.com/aarora79/my-reading-list](https://github.com/aarora79/my-reading-list).
:::
